---
title: "R Practice: Tidy Data and Joins"
format: html
---

# Exercise 1: Practice Joins

## set up
```{r}
#| message: false
#| message: false

library(readr)
library(dplyr)
library(here)
library(lubridate) # for bonus question
```

## read data
```{r}
#| warning: false
#| message: false

bird_observations <- read_csv("https://portal.edirepository.org/nis/dataviewer?packageid=knb-lter-cap.256.10&entityid=53edaa7a0e083013d9bf20322db1780e")
sites <- read_csv("https://portal.edirepository.org/nis/dataviewer?packageid=knb-lter-cap.256.10&entityid=81bf72420e69077097fb0790dcdc63a6")
surveys <- read_csv("https://portal.edirepository.org/nis/dataviewer?packageid=knb-lter-cap.256.10&entityid=b2466fa5cb5ed7ee1ea91398fc291c59")
taxalist <- read_csv("https://portal.edirepository.org/nis/dataviewer?packageid=knb-lter-cap.256.10&entityid=58f863b7e3066e68536a9cacdc7bd58e")
```

## explore data
```{r}
colnames(bird_observations)

head(bird_observations)

summary(bird_observations)

#unique(bird_observations$xx)

glimpse(bird_observations)

head(taxalist)

head(sites)
```

## 10.3 create a subset of bird_observations using 'filter()'

```{r}
birds_subset <- bird_observations %>% 
  filter(species_id %in% c('BHCO', 'RWBL')) %>% 
  filter(site_id %in% c('LI-W', 'NU-C'))
  
head(birds_subset)
```

## 10.4 use 'left_join()' to merge birds_subset with the tables sites
```{r}
birds_left <- birds_subset %>% 
  left_join(y = sites, by = "site_id")

head(birds_left)
```

## 10.5 Use 'full_join()' to merge birds_subset and sites tables
```{r}
birds_full <- birds_subset %>% 
  full_join(y = sites, by = "site_id")

head(birds_full)
```

## Use 'inner_join()' to merge birds_subset and taxalist data
```{r}
birds_inner <- birds_subset %>% 
  inner_join(y = taxalist, by = "species_id")

head(birds_inner)
```

## Use 'left_join()' to merge birds_subset and taxalist data
```{r}
birds_inner <- birds_subset %>% 
  left_join(y = taxalist, by = "species_id")

head(birds_inner)
```


# Exercise 2: Practice Wrangling and Joining Data

# 10.7 Wrangle bird_observations data and merge the data with al the other tables (sites, surveys, and taxalist)

## rename notes to bird_obs_notes
```{r}
bird_observations <- bird_observations %>%
    rename(bird_obs_notes = notes)

head(bird_observations)
head(taxalist)
head(sites)
head(surveys)
```

## Question 7B
```{r}
birds_obs_subset <- bird_observations %>% 
  full_join(y = taxalist, by = "species_id") %>% 
  full_join(y = sites, by = "site_id") %>% 
  full_join(y = surveys, by = c("site_id", "survey_id")) %>% 
  select(survey_date, common_name, park_name, bird_count, observer)

str(birds_obs_subset)

```


## 10.8 Explore observer data and fix the values within this column so that all values are in the same format

```{r}
bird_obs_subset <- birds_obs_subset %>% 
  mutate(observer = if_else(condition = observer == "Josh Burns", 
                            true = "J. Burns", 
                            false = observer))

unique(bird_obs_subset$observer)
```









